---
title: "L2C Smartphone App Paper - Data Import and Cleaning"
date: "2021-08-03 <br> Updated: `r Sys.Date()`"
---


# ⭐️Overview

This file we important and clean data needed for the Link2Care smartphone app paper (still waiting on the title).

2021-07-14, Jordan Neil:
Paper examining prevalence and perceived utility of mHealth tech among RIHA

A call may be helpful, but we might be able to progress with a simple output for both papers.

Can you run a series of zero-order correlations between person-level characteristics and prevalence and perceived utility of mHealth?
I don’t have the list of variable names, but I have attached a table with the list of constructs/items where unclear.
Quick Q: For the item “Type of health related issue managed with smartphone apps” there is a “multiple” response category. Was this from a select all that apply approach that you recoded, or was “multiple” a response item on the survey?


# 📦Load packages

```{r message=FALSE}
library(dplyr, warn.conflicts = FALSE)
library(haven)
library(readr)
```


# 🌎Connect to UTH server 

```{bash eval=FALSE}
# Make sure you are connected to the VPN
open 'smb://islgpcifs.uthouston.edu/sph_research/'
```


# 📥Import data 

The group is only interested in baseline for now. Therefore, I'm importing the QDS visit 1 data only.

```{r}
v1 <- read_sav("/Volumes/sph_research/Link2Care/Participant Data/SPSS Data/raw_data/v1_data.SAV")
```

```{r}
dim(v1) # 324 733
```


# 🚧Data management

Keep variables of interest only

```{r}
v1 <- v1 %>% 
  select(
    id = SUBJECT, gender = GENDER, hispanic = SQ_2, race = SQ_3, age = SQ_4, 
    have_mobile = SQ_12, have_data_plan = SQ_16, edu_20_cat = DEM5V1, 
    employ_9_cat = DEM6V1, lifetime_homeless = BH1V1, lifetime_jail = BH17V1,
    ment_health_treat = BH9V1, app_change = S41_V1, use_app = S42_V1, 
    genhealth = HS1_V1, 
  )
```

## Recode missing

```{r}
v1 <- v1 %>% 
  mutate(
    across(
      .cols = c(app_change, use_app),
      .fns  = ~ ifelse(.x %in% c(7, 8, 9), NA, .x)
    )
  )
```

## Factor variables

```{r}
ny_lab <- c("No", "Yes")
```

```{r}
v1 <- v1 %>% 
  mutate(
    gender_f = factor(gender, 0:2, c("Male", "Female", "Other")),
    hispanic_f = factor(hispanic, 0:1, ny_lab),
    have_mobile_f = factor(have_mobile, 0:1, ny_lab),
    have_data_plan_f = factor(
      have_data_plan, 0:2, c("No", "Yes, limited", "Yes, unlimited")
    ),
    ment_health_treat_f = factor(ment_health_treat, 0:1, ny_lab),
    app_change_f = factor(app_change, 0:1, ny_lab),
    use_app_f = factor(use_app, 0:1, ny_lab),
    genhealth_f = factor(
      genhealth, 1:5, c(
        "Excellent", "Very Good", "Good", "Fair", "Poor"
      )
    )
  )
```

## Recode variables

```{r}
v1 <- v1 %>% 
  mutate(
    
    # Race ethnicity
    race_eth_4_cat = case_when(
      is.na(hispanic) ~ NA_integer_, # Can't categorize at all
      hispanic == 1 ~ 3L, # Hispanic, any race
      is.na(race) ~ NA_integer_, # Non-Hispanic, but we don't know race
      race == 2 ~ 1L, # White, non-Hispanic
      race == 3 ~ 2L, # Black, non-Hispanic
      TRUE ~ 4L # Other race, non-Hispanic
    ),
    race_eth_4_cat_f = factor(
      race_eth_4_cat, 1:4, c(
        "White, non-Hispanic", "Black, non-Hispanic", "Hispanic, any race",
        "Other race, non-Hispanic"
      )
    ),
    
    # Education
    high_school_grad = if_else(edu_20_cat < 12, 0, 1),
    high_school_grad_f = factor(high_school_grad, 0:1, ny_lab),
    
    # Employment status
    employ_5_cat = case_when(
      is.na(employ_9_cat) ~ NA_integer_,
      employ_9_cat %in% c(1, 2) ~ 1L, # Employed full or part time
      employ_9_cat == 3 ~ 2L, # Unemployed, looking for work
      employ_9_cat == 4 ~ 3L, # Unemployed, not looking for work
      employ_9_cat == 8 ~ 4L, # Unable to work, disabled 
      TRUE ~ 5L # Other
    ),
    employ_5_cat_f = factor(
      employ_5_cat, 1:5, c(
        "Employed", "Unemployed, looking for work", 
        "Unemployed, not looking for work", "Unable to work or disabled",
        "Other"
      )
    )
  )
```


# Export analysis data

Export the analysis data as an SPSS data file for Michael and upload it to Kiteworks and the U drive.

We also need to export and rds file to use for analysis in R. Otherwise, the factors are lost.

```{r}
write_sav(
  v1, 
  "/Volumes/sph_research/Link2Care/Statistical Projects/Montgomery - Smartphone/v1_clean.sav"
)
```

```{r}
write_rds(
  v1, 
  "/Volumes/sph_research/Link2Care/Statistical Projects/Montgomery - Smartphone/v1_clean.rds"
)
```


# 🗑Clean up

```{r}
rm(ny_lab)
```

```{r echo=FALSE}
sessionInfo()
```
