---
title: "L2C Smartphone App Paper - Data Import and Cleaning"
date: "2021-08-03 <br> Updated: `r Sys.Date()`"
---


# ⭐️Overview

2022-12-11

Originally, this R project was on Dropbox. On 2022-12-11, we moved it over to OneDrive at "L2C Teams/Abstract and Manuscripts/l2c_paper_smartphone_app". We also modified the file paths below.

2021-08-03

In this file, we important and clean data needed for the Link2Care smartphone app paper (still waiting on the title).

2021-07-14, Jordan Neil:
Paper examining prevalence and perceived utility of mHealth tech among RIHA

A call may be helpful, but we might be able to progress with a simple output for both papers.

Can you run a series of zero-order correlations between person-level characteristics and prevalence and perceived utility of mHealth?
I don’t have the list of variable names, but I have attached a table with the list of constructs/items where unclear.
Quick Q: For the item “Type of health related issue managed with smartphone apps” there is a “multiple” response category. Was this from a select all that apply approach that you recoded, or was “multiple” a response item on the survey?


# 📦Load packages

```{r message=FALSE}
library(dplyr, warn.conflicts = FALSE)
library(haven)
library(readr)
```


# 📥Import data 

The group is only interested in baseline for now. Therefore, I'm importing the QDS visit 1 data only.

```{r}
v1 <- read_sav("../../Participant Data/SPSS Data/raw_data/v1_data.SAV")
```

```{r}
dim(v1) # 442 733
```

Originally, the data we used for this analysis included only 324 rows and 733 columns. We will remove new rows to make the data match.

```{r}
v1 <- slice(v1, 1:324)
```

```{r}
dim(v1) # 324 733
```


# 🚧Data management

Keep variables of interest only

```{r}
v1 <- v1 %>%
  select(
    # Administrative data
    id = SUBJECT, gender = GENDER, 
    
    # Screening questions
    hispanic = SQ_2, race = SQ_3, age = SQ_4, mobile_have = SQ_12, 
    mobile_pays = SQ_13, mobile_talk_min = SQ_14, mobile_smart = SQ_15, 
    mobile_have_data_plan = SQ_16, mobile_number_change = SQ_17, 
    sm_email = SQ_18A, sm_facebook = SQ_18B, sm_google_plus = SQ_18C, 
    sm_twitter = SQ_18D, sm_blogs = SQ_18E, sm_instagram = SQ_18F, 
    sm_snapchat = SQ_18G, sm_linkedin = SQ_18H, sm_none = SQ_18I, 
    access_internet_freq = SQ_19, facebook_freq = SQ_21,
    
    # Demographics questions
    edu_20_cat = DEM5V1, employ_9_cat = DEM6V1, 
    
    # The Brief Homelessness Questionnaire (BH)
    lifetime_homeless = BH1V1, ment_health_treat = BH9V1, lifetime_jail = BH17V1,
    
    # SF-12 Health Survey (HS)
    genhealth = HS1_V1,
    
    # Health Related Quality of Life (HRQ)
    phys_hlth_days = HRQ1_V1, ment_hlth_days = HRQ2_V1,
    
    # Self-Rated Health Questionnaire (S)
    app_change = S41_V1, app_use = S42_V1, app_issues_food = S43_V1A, 
    app_issues_medication = S43_V1B, app_issues_mood = S43_V1C, 
    app_issues_phys_act = S43_V1D, app_issues_sleep = S43_V1E, 
    app_issues_smoking = S43_V1F, app_issues_stress = S43_V1G, 
    app_issues_weight = S43_V1H, app_issues_other = S43_V1I
  )
```

## Recode missing

### Helper function

Convert 9's (or 7's, or 8's) to NA's and then add attributes back to vector.

First, I have to remove the "have_labelled" and "vctrs_vctr" class from each variable for if_else() to work.

But, I also need to keep the label attributes. They are erased when I convert the 9's to NA.

```{r}
nines_to_na <- function(x, nines) {
  # Store the attributes
  x_attr <- attributes(x)
  # Remove "haven_labelled" and "vctrs_vctr" from class list so that if_else()
  # will work.
  # Set class of column to whatever is remaining in the class list.
  classes <- class(x)
  class(x) <- classes[!class(x) %in% c("haven_labelled", "vctrs_vctr")]
  # Convert 9's to NA's
  x <- if_else(x %in% nines, NA_real_, x)
  # Add attributes back to the vector
  attributes(x) <- x_attr
  # Return x
  x
}

# For testing
# test <- all_visits
# test$SQ_13[1] <- 7
# test$SQ_13[3] <- 9
# test$SQ_13[4] <- 99
# test %>% 
#   select(SQ_13) %>% 
#   mutate(SQ_13_test = nines_to_na(SQ_13, c(7, 9, 99))) %>% 
#   pull(SQ_13_test) %>% 
#   attributes()
# rm(test)
```

```{r}
v1 <- v1 %>% 
  mutate(
    across(
      c(app_change, app_use),
      ~ nines_to_na(.x, c(7, 8, 9))
    ),
    across(
      c(mobile_number_change, ),
      ~ nines_to_na(.x, 999)
    )
  )
```

## Factor variables

### Helper function

### Use SPSS labels as factor levels for categorical variables

```{r}
# For use inside across()
spss_to_fs <- function(x) {
  levs <- attr(x, "labels")
  labs <- names(levs)
  x_f <- factor(x, levs, labs)
  x_f
}

# For testing
# all_visits %>%
#   select(SQ_2, SQ_3) %>%
#   mutate(
#     across(
#       everything(),
#       spss_to_fs,
#       .names = "{col}_f"
#     )
#   )
```

### Add SPSS lables to factors

```{r}
# Make factor versions of categorical variables
v1 <- v1 %>% 
  mutate(
    across(
      # It looks like all the categorical variables (e.g., gender) have a class
      # of "haven_labelled". In contrast, the numeric variables (e.g., age) do
      # not
      where(~ "haven_labelled" %in% class(.x)),
      spss_to_fs,
      .names = "{col}_f"
    )
  )
```

## Recode variables

```{r}
v1 <- v1 %>%
  mutate(
    
    # Race ethnicity
    race_eth_4_cat = case_when(
      is.na(hispanic) ~ NA_integer_, # Can't categorize at all
      hispanic == 1 ~ 3L, # Hispanic, any race
      is.na(race) ~ NA_integer_, # Non-Hispanic, but we don't know race
      race == 2 ~ 1L, # White, non-Hispanic
      race == 3 ~ 2L, # Black, non-Hispanic
      TRUE ~ 4L # Other race, non-Hispanic
    ),
    race_eth_4_cat_f = factor(
      race_eth_4_cat, 1:4, c(
        "White, non-Hispanic", "Black, non-Hispanic", "Hispanic, any race",
        "Other race, non-Hispanic"
      )
    ),
    
    # Education
    high_school_grad = if_else(edu_20_cat < 12, 0, 1),
    high_school_grad_f = factor(high_school_grad, 0:1, c("No", "Yes")),
    
    # Employment status
    employ_5_cat = case_when(
      is.na(employ_9_cat) ~ NA_integer_,
      employ_9_cat %in% c(1, 2) ~ 1L, # Employed full or part time
      employ_9_cat == 3 ~ 2L, # Unemployed, looking for work
      employ_9_cat == 4 ~ 3L, # Unemployed, not looking for work
      employ_9_cat == 8 ~ 4L, # Unable to work, disabled 
      TRUE ~ 5L # Other
    ),
    employ_5_cat_f = factor(
      employ_5_cat, 1:5, c(
        "Employed", "Unemployed, looking for work", 
        "Unemployed, not looking for work", "Unable to work or disabled",
        "Other"
      )
    )
  )
```

2022-01-03, from Jordan Neil 

Brad, sorry to ask this again as you provided it for Jillian’s paper, but for this manuscript can you create a category that reports participants number of issues managed by a smartphone app? i.e., we currently have it broken down by issue, but could you also provide how many managed 2 issues, 3 issues, 4 issues, etc.?

NOTE: Person 2296 has a yes for app_issues_other, but NA for app_use. We will not count this (i.e., set to NA).

```{r}
v1 <- v1 %>%
  rowwise() %>% 
  mutate(
    app_issues_total = sum(c_across(app_issues_food:app_issues_other)),
    app_issues_total = if_else(is.na(app_use), NA_real_, app_issues_total),
    app_issues_total_f = factor(app_issues_total, 1:9, 1:9),
    # We are also making a version of this variable that includes zero issues 
    # managed.
    app_issues_total_zero = if_else(app_use == 0, 0, app_issues_total),
    app_issues_total_zero_f = factor(app_issues_total_zero, 0:9, 0:9)
  ) %>% 
  ungroup()
```


# Export analysis data

Export the analysis data as an SPSS data file for Michael and upload it to Kiteworks.

We also need to export and rds file to use for analysis in R. Otherwise, the factors are lost.

```{r}
write_sav(v1, "data/v1_clean.sav")
```

```{r}
write_rds(v1, "data/v1_clean.rds")
```

