---
title: "L2C Smartphone App Paper - Descriptive Table for the Use App question"
date: "2021-08-03 <br> Updated: `r Sys.Date()`"
---


# ⭐️Overview

In this file we create a descriptive table exploring the relationships between using a smartphone app to manage health-related uses and each of the following: sociodemographic background, lifetime homelessness, lifetime incarceration, physical and mental health, and access to a mobile phone and data plan.

Do you believe that a smartphone app can help you to change your actions or behavior?
0	=	No
1	=	Yes
7	=	Don't Know
8	=	Refuse to Answer
9	=	Not Applicable


# 📦Load packages

```{r message=FALSE}
library(dplyr, warn.conflicts = FALSE)
library(freqtables)
library(meantables)
library(purrr)
library(officer)
library(flextable,warn.conflicts = FALSE)
```

```{r}
source("flextable_helpers.R")
```


# 📥Import data 

Import and clean data using data_01_import_clean.Rmd

```{r}
source_rmd <- function(file, ...) {
  tmp_file = tempfile(fileext=".R")
  on.exit(unlink(tmp_file), add = TRUE)
  knitr::purl(file, output=tmp_file)
  source(file = tmp_file, ...)
}
```

```{r}
source_rmd("data_01_import_clean.Rmd")
```

```{r}
dim(v1) # 303 29
```


# 🚧Data management

Drop rows with missing outcome

```{r}
v1_use_app <- v1 %>%
  filter(!is.na(use_app_f))
```


# 📈Analysis

## Outcome distribution

```{r}
v1_use_app %>% 
  freq_table(use_app_f)
```

## Create helper functions

For calculating statistics of interest.

### Categorical statistics

```{r eval=FALSE}
# For testing
v1_use_app %>% 
  freq_table(use_app_f, gender_f) %>% 
  freq_format("percent_row (lcl_row - ucl_row)", digits = 0) %>% 
  select(row_cat, col_var, col_cat, formatted_stats) %>% 
  tidyr::pivot_wider(
    names_from = "row_cat",
    values_from = "formatted_stats"
  ) %>% 
  # Add blank row below
  add_row(col_var = "", col_cat = "", No = "", Yes = "")
```

```{r}
cat_stats_fn <- function(.data, .outcome, .pred, .digits = 0) {
  .data %>% 
    freq_table({{ .outcome }}, {{ .pred }}) %>% 
    freq_format("percent_row (lcl_row - ucl_row)", digits = .digits) %>% 
    select(row_cat, col_var, col_cat, formatted_stats) %>% 
    tidyr::pivot_wider(
      names_from = "row_cat",
      values_from = "formatted_stats"
    ) %>% 
    # Add blank row below
    add_row(col_var = "", col_cat = "", No = "", Yes = "")
}

# For testing
#cat_stats_fn(v1_use_app, use_app_f, gender_f)
```

### Continuous statistics

```{r eval=FALSE}
# For testing
v1_use_app %>% 
  group_by(use_app_f) %>% 
  mean_table(age) %>% 
  freq_format("mean (lcl - ucl)", digits = 0) %>% 
  select(response_var, group_cat, formatted_stats) %>% 
  tidyr::pivot_wider(
    names_from = "group_cat",
    values_from = "formatted_stats"
  ) %>% 
  # Add blank row below
  add_row(response_var = "", No = "", Yes = "")
```

```{r}
cont_stats_fn <- function(.data, .outcome, .pred, .digits = 0) {
  .data %>% 
    group_by({{ .outcome }}) %>% 
    mean_table({{ .pred }}) %>% 
    freq_format("mean (lcl - ucl)", digits = .digits) %>% 
    select(response_var, group_cat, formatted_stats) %>% 
    tidyr::pivot_wider(
      names_from = "group_cat",
      values_from = "formatted_stats"
    ) %>% 
    # Add blank row below
    add_row(response_var = "", No = "", Yes = "")
}

# For testing
# cont_stats_fn(v1_use_app, use_app_f, age)
```

```{r}
cat_cols <- quos(
  gender_f, race_eth_4_cat_f, high_school_grad_f, employ_5_cat_f, genhealth_f,
  ment_health_treat_f, have_mobile_f, have_data_plan_f, 
)
```

```{r}
cont_cols <- quos(age, lifetime_homeless, lifetime_jail)
```

```{r}
map_df(
  .x = cat_cols,
  .f = ~ cat_stats_fn(v1_use_app, use_app_f, !! .x)
)
```
```{r}
table(v1$have_data_plan)
```


# Left off here. Have data plan has a ton of missing.


```{r}
# Loop over all categorical vars
# t1_cat_stats_no <- 
map_df(
  cat_cols, 
  function(x) {
    v1 %>%
      filter(screened_in == 1) %>% 
      freq_table({{x}}) %>%
      freq_format(recipe = "n (percent)", digits = 1) %>%
      select(var, cat, formatted_stats) %>%
      # Add a row with the var name only
      add_row(var = quo_name(x), .before = 1) %>% 
      # Add blank row below
      add_row(var = "", cat = "", formatted_stats = "")
  }
) %>% 
  # Drop the final empty row
  slice(-n())
```




# 🗑Clean up

```{r}
rm(list = ls())
```

```{r echo=FALSE}
sessionInfo()
```
