---
title: "L2C Smartphone App Paper - Descriptive Table for the Use App question"
date: "2021-08-03 <br> Updated: `r Sys.Date()`"
---

# Instructions

Make any data updates you need to make in data_01_import_clean.Rmd. After that file has saved a new version of "/Volumes/sph_research/Link2Care/Statistical Projects/Montgomery - Smartphone/v1_clean.rds", you should just be able to click Run > Run All Chunks Below. This will create an updated Word document in the root directory called, "Smartphone Table Updates.docx".


# ⭐️Overview

In this file we create a descriptive table exploring the relationships between using a smartphone app to manage health-related uses and each of the following: sociodemographic background, lifetime homelessness, lifetime incarceration, physical and mental health, and access to a mobile phone and data plan.

The two main outcomes of interest are:

Do you believe that a smartphone app can help you to change your actions or behavior? (v1_app_change)
0	=	No
1	=	Yes
7	=	Don't Know
8	=	Refuse to Answer
9	=	Not Applicable


Have you ever used a smartphone app to manage one or more health-related issues? (v1_use_app)
0	=	No
1	=	Yes
7	=	Don't Know
8	=	Refuse to Answer
9	=	Not Applicable


# 📦Load packages

```{r message=FALSE}
library(dplyr, warn.conflicts = FALSE)
library(readr)
library(freqtables)
library(meantables)
library(purrr)
library(officer)
library(flextable, warn.conflicts = FALSE)
```

```{r}
source("flextable_helpers.R")
```


# 🌎Connect to UTH server 

```{bash eval=FALSE}
# Make sure you are connected to the VPN
open 'smb://islgpcifs.uthouston.edu/sph_research/'
```


# 📥Import data 

This data is created and saved in data_01_import_clean.Rmd

```{r}
v1 <- read_rds("/Volumes/sph_research/Link2Care/Statistical Projects/Montgomery - Smartphone/v1_clean.rds")
```

```{r}
dim(v1) # 324  82
```


# 📈Analysis

2021-12-01: Jordan and Michael no longer want Table 1 broken down by app_change (Yes/NO) and they no longer want table 2 broken down by app_use (Yes/No).

## Create helper functions

For calculating statistics of interest.

### Categorical statistics

```{r eval=FALSE}
# For testing
v1 %>% 
  filter(!is.na(app_use_f)) %>% 
  freq_table(app_use_f, drop = TRUE) %>% 
  freq_format("n (percent)", digits = 1) %>% 
  select(var, cat, formatted_stats) %>% 
  # Add two spaces in front of categories. This is to reduce the amount of 
  # formatting we have to do in Word. The spaces won't be visible while testing
  # in RStudio, but they will show up in the Word document
  mutate(cat = paste0("  ", cat)) %>% 
  # Add top row to contain the var name in the cat column. This is to 
  # reduce the amount of formatting we have to do in Word. 
  add_row(cat = .$var[1], .before = 1) %>% 
  # Now drop var
  select(-var, var = cat) %>% 
  # Add blank row below
  add_row(var = "", formatted_stats = "")
```

```{r}
cat_stats_fn <- function(.data, .outcome, .digits = 0) {
  .data %>% 
    filter(!is.na({{ .outcome }})) %>% 
    freq_table({{ .outcome }}, drop = TRUE) %>% 
    freq_format("n (percent)", digits = .digits) %>% 
    select(var, cat, formatted_stats) %>% 
    # Add two spaces in front of categories. This is to reduce the amount of 
    # formatting we have to do in Word. The spaces won't be visible while testing
    # in RStudio, but they will show up in the Word document
    mutate(cat = paste0("  ", cat)) %>%
    # Add top row to contain the var name in the col_cat column. This is to 
    # reduce the amount of formatting we have to do in Word. 
    add_row(cat = .$var[1], .before = 1) %>% 
    # Now drop var
    select(-var, var = cat) %>% 
    # Add blank row below
    add_row(var = "", formatted_stats = "")
}

# For testing
# cat_stats_fn(v1, app_issues_other_f, .digits = 1)
# cat_stats_fn(v1, !!sym("gender_f"))
```

### Continuous statistics

```{r eval=FALSE}
# For testing
v1 %>% 
  filter(!is.na(age)) %>% 
  mean_table(age) %>% 
  mean_format("mean (sd)", digits = 1) %>% 
  select(var = response_var, formatted_stats) %>% 
  # Add blank row below
  add_row(var = "", formatted_stats = "")
```

```{r}
cont_stats_fn <- function(.data, .outcome, .digits = 0) {
  .data %>% 
    filter(!is.na({{ .outcome }})) %>%
    mean_table({{ .outcome }}) %>%
    mean_format("mean (sd)", digits = .digits) %>% 
    select(var = response_var, formatted_stats) %>% 
    # Add blank row below
    add_row(var = "", formatted_stats = "")
}

# For testing
# cont_stats_fn(v1, age, .digits = 1)
```

### Dummy variables

The "Which of the following forms of media do you use?" and the "What type of health related issue?" questions both allow multiple answers. This results in dummy variables for each possible response option. For example:

|------------|:--------:|
| sm_email_f | NA       |
| No         | 74 (23)  |
| Yes        | 250 (77) |

Instead, we want to treat each of the "yes" rows for each of the individual dummy variables as though they were categories of a higher level variable. For example:

|--------------|:--------:|
| Use media    | NA       |
|   Email      | 250 (77) |
|   Twitter    | 29 (9)   |

```{r eval=FALSE}
# For testing
v1 %>% 
  filter(!is.na(sm_email_f)) %>% 
  freq_table(sm_email_f, drop = TRUE) %>% 
  freq_format("n (percent)", digits = 1) %>% 
  # Only keep the "Yes" row
  filter(cat == "Yes") %>% 
  select(var, formatted_stats) %>% 
  # Add two spaces in front of categories. This is to reduce the amount of 
  # formatting we have to do in Word. The spaces won't be visible while testing
  # in RStudio, but they will show up in the Word document
  mutate(var = paste0("  ", var))
```

```{r}
# Function to calculate the stats of interest for a single dummy variable
dummy_stats_fn <- function(.data, .outcome, .digits = 0) {
  .data %>% 
    filter(!is.na({{ .outcome }})) %>% 
    freq_table({{ .outcome }}, drop = TRUE) %>% 
    freq_format("n (percent)", digits = .digits) %>% 
    # Only keep the "Yes" row
    filter(cat == "Yes") %>% 
    select(var, formatted_stats) %>% 
    # Add two spaces in front of categories. This is to reduce the amount of 
    # formatting we have to do in Word. The spaces won't be visible while testing
    # in RStudio, but they will show up in the Word document
    mutate(var = paste0("  ", var))
}

# For testing
# dummy_stats_fn(v1, sm_email_f, .digits = 1)
```

Apply dummy_stats_fn to multiple dummy variables and bind the results together as though they were categories of a single variable. 

```{r}
# For testing
map_df(
  .x = c("sm_email_f", "sm_facebook_f"),
  .f = ~ dummy_stats_fn(v1, !! sym(.x))
) %>% 
  # Add text indicating the name of the higher-level parent question to the first
  # row of the data frame
  add_row(var = "media_types", formatted_stats = NA, .before = 1) %>% 
  # Add blank row below
  add_row(var = "", formatted_stats = "")
```

```{r}
dummies_stats_fn <- function(.data, .parent_q, .dummies, .digits = 0) {
  map_df(
    .x = .dummies,
    .f = ~ dummy_stats_fn(.data, !! sym(.x), .digits = .digits)
  ) %>% 
    # Add text indicating the name of the higher-level parent question to the first
    # row of the data frame
    add_row(var = .parent_q, formatted_stats = NA, .before = 1) %>% 
    # Add blank row below
    add_row(var = "", formatted_stats = "")
}

# For testing
# dummies_stats_fn(v1, "media_types", c("sm_email_f", "sm_facebook_f"), .digits = 1)
```

## Create data frames of stats for tables

```{r}
# For checking
# select(v1, ends_with("_f")) %>% names()
# select(v1, where(~ !("haven_labelled" %in% class(.x)) && !is.factor(.x)))
```

```{r}
cat_cols <- c(
  "gender_f", "race_eth_4_cat_f", "high_school_grad_f", "employ_5_cat_f", 
  "genhealth_f", "ment_health_treat_f", "mobile_have_f", "mobile_pays_f",
  "mobile_talk_min_f", "mobile_smart_f", "mobile_have_data_plan_f", 
  "mobile_number_change_f", "access_internet_freq_f", "facebook_freq_f", 
  "app_change_f", "app_use_f"
)
```

```{r}
cont_cols <- c(
  "age", "lifetime_homeless", "lifetime_jail", "phys_hlth_days", "ment_hlth_days"
)
```

```{r}
dummy_cols <- list(
  # Which of the following forms of media do you use?
  media_types = c(
  "sm_email_f", "sm_facebook_f", "sm_google_plus_f", "sm_twitter_f", 
  "sm_blogs_f","sm_instagram_f", "sm_snapchat_f", "sm_linkedin_f", "sm_none_f"
  ),
  
  # What type of health related issue? [use smartphone app to manage]
  app_health_issue = c(
  "app_issues_food_f", "app_issues_medication_f", "app_issues_mood_f", 
  "app_issues_phys_act_f", "app_issues_sleep_f", "app_issues_smoking_f", 
  "app_issues_stress_f", "app_issues_weight_f", "app_issues_other_f"
  )
)
```

### Create lists of statistics of interest

Calculate categorical stats

```{r}
stats_list <- cat_cols %>%
  set_names(cat_cols) %>% 
  map(~ cat_stats_fn(v1, !! sym(.x), .digits = 1))
```

Calculate continuous stats

```{r}
stats_list <- c(
  stats_list,
  cont_cols %>%
    set_names(cont_cols) %>%
    map(~ cont_stats_fn(v1, !! sym(.x), .digits = 1))
)
```

Calculate categorical stats for dummy variables

```{r}
stats_list <- c(
  stats_list,
  imap(
    dummy_cols,
    ~ dummies_stats_fn(v1, .y, .x, .digits = 1)
  )
)
```


#### Bind together categorical and continuous stats

```{r}
# Use this later to reorder the variables in the tables
variable_order <- c(
  # Sociodemographics
  "age", "gender_f", "race_eth_4_cat_f", "high_school_grad_f", "employ_5_cat_f",
  "lifetime_homeless", "lifetime_jail",
  
  # Health
  "genhealth_f", "phys_hlth_days", "ment_hlth_days", "ment_health_treat_f",
  
  # Technology use
  "mobile_have_f", "mobile_pays_f",
  "mobile_talk_min_f", "mobile_smart_f", "mobile_have_data_plan_f", 
  "mobile_number_change_f", "media_types", "access_internet_freq_f", 
  "facebook_freq_f", "app_change_f", "app_use_f", "app_health_issue"
)
```

```{r}
table_1 <- map_dfr(
  .x = variable_order,
  .f = ~ bind_rows(stats_list[[.x]])
)
```

Remove the final blank row. It makes the Word table look cleaner.

```{r}
table_1 <- slice(table_1, -nrow(table_1))
```

# Format tables

To reduce the amount of formatting that needs to be done in word. 

### Make variable names easier for end users to read

We want to make variable names easier for end users to read and also add the statistics used to the row headers.

```{r}
change_row_names <- function(.data) {
  .data %>% 
    mutate(var = case_when(
      # Sociodemographics
      var == "age" ~ "Age, mean (sd)",
      var == "gender_f" ~ "Gender, n (percent)",
      var == "race_eth_4_cat_f" ~ "Race/Ethnicity, n (percent)",
      var == "high_school_grad_f" ~ "High school grad or GED, n (percent)",
      var == "employ_5_cat_f" ~ "Employment status, n (percent)",
      var == "lifetime_homeless" ~ "Lifetime months homeless, mean (sd)",
      var == "lifetime_jail" ~ "Lifetime years in jail, mean (sd)",
      
      # Health
      var == "genhealth_f" ~ "General health, n (percent)",
      var == "phys_hlth_days" ~ "N days out of past 30 physical health not good, mean (sd)", 
      var == "ment_hlth_days" ~ "N days out of past 30 mental health not good, mean (sd)",
      var == "ment_health_treat_f" ~ "Mental health treatment, n (percent)",
      
      # Technology use
      var == "mobile_have_f" ~ "Have mobile phone, n (percent)",
      var == "mobile_pays_f" ~ "Mobile phone bill payer, n (percent)",
      var == "mobile_talk_min_f" ~ "Talk minutes in mobile plan, n (percent)", 
      var == "mobile_smart_f" ~ "Is mobile phone a smart phone, n (percent)",
      var == "mobile_have_data_plan_f" ~ "Have data plan, n (percent)",
      var == "mobile_number_change_f" ~ "N times mobile number has changed, n (percent)", 
      var == "media_types" ~ "Types of media used, n (percent)",
      var == "  sm_email_f" ~ "  Email", 
      var == "  sm_facebook_f" ~ "  Facebook", 
      var == "  sm_google_plus_f" ~ "  Google Plus",
      var == "  sm_twitter_f" ~ "  Twitter",
      var == "  sm_blogs_f" ~ "  Blogs",
      var == "  sm_instagram_f" ~ "  Instagram",
      var == "  sm_snapchat_f" ~ "  Snapchat",
      var == "  sm_linkedin_f" ~ "  LinkedIn",
      var == "  sm_none_f" ~ "  None",
      var == "access_internet_freq_f" ~ "Frequency of internet access, n (percent)",
      var == "facebook_freq_f" ~ "Frequency of Facebook use, n (percent)",
      var == "app_change_f" ~ "Believe smartphone app can help change actions or behaviors, n (percent)",
      var == "app_use_f" ~ "used smartphone app to manage health-related issues, n (percent)",
      var == "app_health_issue" ~ "Type of issue managed with smartphone app, n (percent)",
      var == "  app_issues_food_f" ~ "  Food or calorie tracking",
      var == "  app_issues_medication_f" ~ "  Medication reminders",
      var == "  app_issues_mood_f" ~ "  Mood manager",
      var == "  app_issues_phys_act_f" ~ "  Physical activity",
      var == "  app_issues_sleep_f" ~ "  Sleep tracker",
      var == "  app_issues_smoking_f" ~ "  Smoking Cessation",
      var == "  app_issues_stress_f" ~ "  Stress reduction",
      var == "  app_issues_weight_f" ~ "  Weight loss tracking",
      var == "  app_issues_other_f" ~ "  Other",
      TRUE ~ var
    ))
}

# For testing
# change_row_names(table_1)
```

```{r}
table_1 <- change_row_names(table_1)
```


# Create flextables

All the stats are stored in data frames. Next, we will convert those data frames to flextables that we can add to our Word document template.

## Make column headers

```{r}
table_1_ft <- flextable(table_1) %>%
  # Change column widths. figure out through trial and error
  width(width = c(4.69, 1.81)) %>%
  # Center the final two columns
  align(j = 2, align = "center", part = "all") %>% 
  # Change header names -- add subgroup Ns to headers
  set_header_labels(
    var = "Characteristic",
    formatted_stats = "Statistics"
  ) %>% 
  # Bold column headers
  bold(part = "header") %>% 
  footnote(
    i = c(67, 80, 119), j = 1,
    value = as_paragraph(c(
      "Have data plan was only asked of participants who reported having a mobile phone.",
      "Percentages sum to >100% because participants could select more than one response option.",
      "Percentages sum to >100% because participants could select more than one response option."
    ))
  ) %>% 
  my_ft_theme()
```

# 📝Load Word template for officer

```{r}
doc <- read_docx("template_descriptive_analysis.docx") %>%
  body_replace_text_at_bkm("date", as.character(Sys.Date())) %>% 
  body_replace_text_at_bkm("n_baseline", as.character(nrow(v1))) %>% 
  body_replace_flextable_at_bkm("table_baseline_characteristics", table_1_ft) 
```


# 📌Generate Word reports

```{r}
print(
  doc,
  "Smartphone Table Updates.docx"
)
```

